---
draft: true
title: Working with EIA Data
author: Gregory Vitz
date: '2019-05-29'
slug: working-with-eia-data
categories: []
tags: []
Categories:
  - r
Description: ''
Tags:
  - r
  - eia
menu: ''
---

The United States Energy Information Administration (EIA) offers thousands of free data sets publicly, and they can all be found on their website [here](https://www.eia.gov/electricity/data/eia860/). I believe that each of the datasets can be downloaded as a CSV, however there is also an option to access the data through the EIA API. Today, I will be working with some data that I manually downloaded from the EIA website with generator level information. A future post may address how to use the EIA API to programatically get the data. 

```{r message=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Using `readxl` we can easily import the data downloaded from EIA. I used two separate pipes to handle some extra notes added at the bottom of the 

```{r}
file <- here("static", "data", "eia8602018er", "3_1_Generator_Y2018_Early_Release.xlsx")

operable <- "Operable" %>%  
    set_names() %>% 
    read_excel(skip = 2, path = file, guess_max = 20000, n_max = 22050)

proposed <- "Proposed" %>%  
    set_names() %>% 
    read_excel(skip = 2, path = file, guess_max = 2000, n_max = 1218)

```



