---
draft: false
title: Working with EIA Data
author: Gregory Vitz
date: '2019-05-29'
slug: working-with-eia-data
categories: []
tags: []
Categories:
  - r
Description: ''
Tags:
  - r
  - eia
menu: ''
---

The United States Energy Information Administration (EIA) offers thousands of free data sets publicly, and they can all be found on their website [here](https://www.eia.gov/electricity/data/eia860/). I believe that each of the datasets can be downloaded as a CSV, however there is also an option to access the data through the EIA API. Today, I will be working with some data that I manually downloaded from the EIA website with generator level information. A future post may address how to use the EIA API to programatically get the data. 

```{r message=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Using `readxl` we can easily import the data downloaded from EIA. I used two separate pipes to handle some extra notes added at the bottom of the 

```{r}
file <- here("static", "data", "eia8602018er", "3_1_Generator_Y2018_Early_Release.xlsx")
plant_file <- here("static", "data", "eia8602018er", "2___Plant_Y2018_Early_Release.xlsx")


operable <- "Operable" %>%  
    set_names() %>% 
    read_excel(skip = 2, path = file, guess_max = 20000, n_max = 22050)

retired <- "Retired and Canceled" %>%  
    set_names() %>% 
    read_excel(skip = 2, path = file, guess_max = 4000, n_max = 4324)

plant <- "Plant" %>% 
  set_names() %>% 
  read_excel(skip = 2, path = plant_file, guess_max = 4000, n_max = 4324) %>% 
  select(`Plant Code`, `Balancing Authority Code`)

glimpse(operable)
```

As you can see there are a ton of variables in this data. Since we plan on doing an analysis of added and lost capacity we will only need a few.

```{r}
# Units that began operation in 2018
operable2018 <- operable %>% 
    filter(`Operating Year` == 2018) %>%
    transmute(`Plant Code`, 
              `Nameplate Capacity (MW)`, 
              `Energy Source 1`,
              status = 1)

# Units that retired in 2018
retired2018 <- retired %>% 
    filter(`Retirement Year` == 2018) %>% 
    transmute(`Plant Code`, 
              `Nameplate Capacity (MW)`, 
              `Energy Source 1`,
              status = 0)

all_data <- bind_rows(operable2018, retired2018) %>% 
    left_join(plant, by = "Plant Code") %>% 
    rename(plant_code = `Plant Code`, 
           nameplate = `Nameplate Capacity (MW)`, 
           energy_source = `Energy Source 1`,
           ba = `Balancing Authority Code`)
```

Joining the plant capacity and fuel data with balancing authority information will give an idea of what regions were impacted most by additions and retirements. Some minor data manipulation is needed in order to create a stacked bar chart. 

```{r}
all_data <- all_data %>% 
    mutate(capacity = if_else(status == 1, 
                              nameplate, 
                              -nameplate))
```

Creating the first ggplot

```{r}
all_data %>% 
    ggplot(aes(ba, capacity, fill = energy_source)) +
    geom_bar(stat = "identity", position = "stack")
```

Not very pretty. Some easy changes could be made to improve the plot quickly. Let's set a universal theme and reclassify some of the energy sources, while focusing on the major ISO/RTO regions in the United States.

```{r}
theme_set(theme_minimal())

coal <- c("ANT", "BIT", "LIG", "SGC", "SUB", "WC", "RC")
petrol <- c("DFO", "JF", "KER", "PC", "PG", "RFO", "SGP", "WO")
other_gas <- c("BFG",  "OG")
other_renew <- c("OBL", "SLW", "BLQ", "WDL", "LFG", "OBG")


all_data2 <- all_data %>% 
    mutate(source_recode = case_when(
        energy_source %in% coal ~ "Coal",
        energy_source %in% petrol ~ "Petroleum Products",
        energy_source %in% other_gas ~ "Other Gas",
        energy_source %in% other_renew ~ "Other Renewables",
        energy_source == "NG" ~ "Natural Gas",
        energy_source == "SUN" ~ "Solar",
        energy_source == "WND" ~ "Wind",
        energy_source == "GEO" ~ "Geothermal",
        energy_source == "WAT" ~ "Hydro",
        energy_source == "NUC" ~ "Nuclear",
        TRUE ~ "Other"
    )) 

all_data2 %>% 
    ggplot(aes(fct_reorder(ba, capacity, sum), capacity, fill = source_recode)) +
    geom_bar(stat = "identity", position = "stack") +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
```

Lots of capacity was added in regions that we don't have BA codes for. If we drop those areas, we will get a better idea of which Balancing areas had capacity additions/losses.

```{r}
all_data2 %>% 
    filter(ba != "NA") %>%
    ggplot(aes(fct_reorder(ba, capacity, sum), capacity, fill = source_recode)) +
    geom_bar(stat = "identity", position = "stack") +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    scale_y_continuous(labels = scales::comma)
```

